The Heavy Photon Search Test Detector

### 2014-03-06 hps-testrun-nim.tex
Updating mip, eoverp and turn on plot to latest versions

These hits are passed through a simple clustering algorithm which forms clusters by grouping adjacent 
strips. The position of a cluster on the sensor is determined by the amplitude-weighted mean.
With a linear gain up to $\approx 3$~MIPs, the cluster charge for hits associated with a track follow 
the characteristic Landau shape as expected, see Fig.~\ref{fig:cluster_pulse}.  
A noise level between $1.1-1.5\times 10^{3}$ electrons was established through multiple calibration 
runs giving a signal to noise ratio of $21-25$. Lab-based radioactive source tests was used to 
provide the absolute charge normalization.
\begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/run1351_mip_small.pdf}
    	\caption{ The cluster charge distribution for hits associated with a track. }
	\label{fig:cluster_pulse}
}
\end{center}
\end{figure}

After clustering hits on a sensor, the hit time for each cluster is computed as the amplitude-weighted 
average of the fitted $t_0$ channel times. The $t_0$-resolution is studied by comparing the cluster hit 
time with the average of all cluster hit times on the track, the ``track time'', which has the expected jitter 
due to clock phase and trigger, approximately 25~ns. Figure~\ref{fig:tracktime} shows the residual to the individual cluster 
times. 
\begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=6cm]{figures/test_run_1351_hit_time_corrected_top_layer2.pdf}
	\caption{
	The cluster time residual for a representative sensor relative to the track time (see text for more 
	details). }
	\label{fig:tracktime}
}
\end{center}
\end{figure}
After correcting for offsets from each sensor (time-of-flight, clock phase) and accounting for the 
correlation between the $t_0$ and track time,  the extracted $t_0$ resolution is 2.6~ns. This is somewhat 
worse than the approximately 2~ns resolution expected which we attribute to the true 
pulse shape differing from our idealized fit function which will be improved in the future. Reducing the 
APV25 pulse shaping time will also improve time resolution. These results show that we can operate 
with the six sample readout mode of the APV25 chip and achieve time resolution adequate for pileup 
rejection during electron running in the HPS experiment.

@@

The ratio of the ECal cluster energy $E$ to the momentum $p$ of a matched track in the SVT was used 
to determine the conversion factors from ADC counts to energy. To compare data and simulation, all 
inoperable or noisy channels in the SVT and ECal was disabled in both data and simulation; any 
efficiency or bias that affect the data should be reflected in the simulation. 
%Furthermore, a conversion factor calibration is performed to equalize the gains of  each channel.  simulation to those observed in data by comparing .  The ``weighted $E/p$'' for each crystal is computed, 
representing the average $E/p$ for clusters that include the crystal
% $\frac{\sum_j w_{j,i}}{\sum_j\frac{P_j}{E_j}w_{j,i}}$.
Iteratively, conversion coefficients for each module is adjusted until the $E/p$ ratio in data and 
simulation are similar. The distribution of the $E/p$ ratio in data and simulation can be seen in 
Fig.~\ref{fig:gains}. 
{\small
\begin{figure}[]
\begin{center}
	\includegraphics[width=7cm]{figures/h_ep_data_h_ep_mc_dataMC_v5_1351_goodregion_notrkcuts.pdf}
	\caption{The ECal energy over track momentum ratio comparing data and simulation} 
	\label{fig:gains}
\end{center}
\end{figure}
}
The peak position of the distribution indicates the sampling fraction of the ECal, 
the fraction of the incident particle energy measured in the cluster. The width and tails of the distribution 
in data indicates imperfect calibration and noise of the ECal modules. This level of calibration and the 
agreement with simulation was found to be sufficient for the multiple scattering analysis using 
normalized event rates described in Sec.~\ref{sec:mcs}.

@@

\subsection{Trigger Performance}
The trigger and DAQ integrate pulses from the ECal modules differently to measure the energy 
deposited in the each crystal. The trigger integrates the pulse using a time-over-threshold window, and 
the DAQ readout integrates using a fixed-size window (5 samples before and 30 samples after a 
threshold crossing). For every event, the trigger reports the trigger decision as a bit mask (top half, 
bottom half or both) and the time the trigger fired. The trigger performance was studied by simulating the 
trigger for each event and comparing to how the events were actually triggered.
First, we simulate the FADC readout board to convert from readout hits (with fixed-size window 
integration) to trigger hits (time-over-threshold integration). Secondly, the CTP clustering 
algorithm and the trigger decision is simulated before we compare the trigger decision and trigger time 
to what was reported by the actual trigger. To eliminate trigger bias, we use a tag and probe method. To 
study the trigger performance in one half of the ECal, we select events which triggered the other half 
and where there was exactly one probe cluster in the ECal half under study. We then measure trigger 
efficiency as the fraction of tagged events that fired the trigger in the probe half as a function of the 
probe cluster energy, shown in Fig.~\ref{fig:turnon}. As a cross-check, the turn-on from the trigger 
threshold was measured to be 1280 in units of ADC counts as expected. The threshold was not 
perfectly sharp because of uncertainties in the conversion from readout to trigger hits described above, 
but based on comparisons with simulation we found that the trigger worked exactly as specified.
\begin{figure}[ht]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/top_turn_on__bottom_tag_E}
	\caption{Trigger turn-on in the top half of the ECal (bottom is similar) for data and simulation as a 
	function of probe cluster energy.}
\label{fig:turnon}
}
\end{center}
\end{figure}
The trigger turn-on is slow and reaches an uneven plateau just below $1$ GeV for two reasons;  gain 
variations between different crystals lead to the threshold variations and the nonlinearity of the 
time-over-threshold integral means that the effective threshold is higher for clusters that span multiple 
crystals. The effective trigger threshold is therefore dependent on position and energy of the particle 
as well as cluster multiplicity. 
%Overall the trigger appears to have functioned exactly as intended. 
%Changes planned for physics running (constant integration window and per-crystal gain calibration 
%constants for the trigger) will solve both of the issues that led to threshold variations in the Test run.

>>>>

These hits are passed through a simple clustering algorithm which forms clusters by grouping adjacent 
strips. The position of a cluster on the sensor is determined by the amplitude-weighted mean.
With a linear gain up to $\approx 3$~MIPs, the cluster charge for hits associated with a track follow 
the characteristic Landau shape as expected, see Fig.~\ref{fig:cluster_pulse}.  
A noise level between $1.1-1.5\times 10^{3}$ electrons was established through multiple calibration 
runs giving a signal to noise ratio of $21-25$. Lab-based radioactive source tests was used to 
provide the absolute charge normalization.
\begin{figure}[]
\begin{center}
{\small
	%\includegraphics[width=7cm]{figures/run1351_mip_small.pdf}
	\includegraphics[width=7cm]{figures/mip_top_layer_2.png}
    	\caption{ The cluster charge distribution for hits associated with a track. }
	\label{fig:cluster_pulse}
}
\end{center}
\end{figure}

After clustering hits on a sensor, the hit time for each cluster is computed as the amplitude-weighted 
average of the fitted $t_0$ channel times. The $t_0$-resolution is studied by comparing the cluster hit 
time with the average of all cluster hit times on the track, the ``track time'', which has the expected jitter 
due to clock phase and trigger, approximately 25~ns. Figure~\ref{fig:tracktime} shows the residual to the individual cluster 
times. 
\begin{figure}[]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/test_run_1351_hit_time_corrected_top_layer2_mod.pdf}
	\caption{
	The cluster time residual for a representative sensor relative to the track time (see text for more 
	details). }
	\label{fig:tracktime}
}
\end{center}
\end{figure}
After correcting for offsets from each sensor (time-of-flight, clock phase) and accounting for the 
correlation between the $t_0$ and track time,  the extracted $t_0$ resolution is 2.6~ns. This is somewhat 
worse than the approximately 2~ns resolution expected which we attribute to the true 
pulse shape differing from our idealized fit function which will be improved in the future. Reducing the 
APV25 pulse shaping time will also improve time resolution. These results show that we can operate 
with the six sample readout mode of the APV25 chip and achieve time resolution adequate for pileup 
rejection during electron running in the HPS experiment.

@@

The ratio of the ECal cluster energy $E$ to the momentum $p$ of a matched track in the SVT was used 
to determine the conversion factors from ADC counts to energy. To compare data and simulation, all 
inoperable or noisy channels in the SVT and ECal was disabled in both data and simulation; any 
efficiency or bias that affect the data should be reflected in the simulation. 
%Furthermore, a conversion factor calibration is performed to equalize the gains of  each channel.  simulation to those observed in data by comparing .  The ``weighted $E/p$'' for each crystal is computed, 
representing the average $E/p$ for clusters that include the crystal
% $\frac{\sum_j w_{j,i}}{\sum_j\frac{P_j}{E_j}w_{j,i}}$.
Iteratively, conversion coefficients for each module is adjusted until the $E/p$ ratio in data and 
simulation are similar. The distribution of the $E/p$ ratio in data and simulation can be seen in 
Fig.~\ref{fig:gains}. 
{\small
\begin{figure}[]
\begin{center}
	\includegraphics[width=7cm]{figures/h_ep_data_0_h_ep_MC_0_dataMC_1351-v6-v6gains_2-trig-top-cl600reg0.pdf}
	\caption{The ECal energy over track momentum ratio comparing data and simulation} 
	\label{fig:gains}
\end{center}
\end{figure}
}
The peak position of the distribution indicates the sampling fraction of the ECal, 
the fraction of the incident particle energy measured in the cluster. The width and tails of the distribution 
in data indicates imperfect calibration and noise of the ECal modules. This level of calibration and the 
agreement with simulation was found to be sufficient for the multiple scattering analysis using 
normalized event rates described in Sec.~\ref{sec:mcs}.

@@

\subsection{Trigger Performance}
The trigger and DAQ integrate pulses from the ECal modules differently to measure the energy 
deposited in the each crystal. The trigger integrates the pulse using a time-over-threshold window, and 
the DAQ readout integrates using a fixed-size window (5 samples before and 30 samples after a 
threshold crossing). For every event, the trigger reports the trigger decision as a bit mask (top half, 
bottom half or both) and the time the trigger fired. The trigger performance was studied by simulating the 
trigger for each event and comparing to how the events were actually triggered.
First, we simulate the FADC readout board to convert from readout hits (with fixed-size window 
integration) to trigger hits (time-over-threshold integration). Secondly, the CTP clustering 
algorithm and the trigger decision is simulated before we compare the trigger decision and trigger time 
to what was reported by the actual trigger. To eliminate trigger bias, we use a tag and probe method. To 
study the trigger performance in one half of the ECal, we select events which triggered the other half 
and where there was exactly one probe cluster in the ECal half under study. We then measure trigger 
efficiency as the fraction of tagged events that fired the trigger in the probe half as a function of the 
probe cluster energy, shown in Fig.~\ref{fig:turnon}. As a cross-check, the turn-on from the trigger 
threshold was measured to be 1280 in units of ADC counts as expected. The threshold was not 
perfectly sharp because of uncertainties in the conversion from readout to trigger hits described 
above, but based on comparisons with simulation we found that the trigger worked exactly as 
specified.
\begin{figure}[ht]
\begin{center}
{\small
	\includegraphics[width=7cm]{figures/h_cl_E_probedata_eff_h_cl_E_probeMC_eff_dataMC_1351-v7-trig-tagBot.pdf}
	\caption{Trigger efficiency in both halves of the ECal for data and simulation as a 
	function of cluster energy.}
\label{fig:turnon}
}
\end{center}
\end{figure}
The trigger turn-on is slow and reaches an uneven plateau just below 1~GeV for two reasons;  
gain variations between different crystals lead to the threshold variations and the nonlinearity of 
the time-over-threshold integral means that the effective threshold is higher for clusters that span 
multiple crystals. The effective trigger threshold is therefore dependent on position and energy of 
the particle as well as cluster multiplicity. 
%Overall the trigger appears to have functioned exactly as intended. 
%Changes planned for physics running (constant integration window and per-crystal gain calibration 
%constants for the trigger) will solve both of the issues that led to threshold variations in the Test run.
